---
title: "HW3"
output:
  pdf_document: default
  html_document: default
---

```{r, echo=FALSE, warning=FALSE}
# Importing packages
library(tidyverse)

library(mosaic)

# Reading datasets into variables
gas_prices <- read.csv("C:\\Users\\gelet\\OneDrive\\Documents\\SDS315\\HW3\\gasprices.csv")

s_class <- read.csv("C:\\Users\\gelet\\OneDrive\\Documents\\SDS315\\HW3\\sclass.csv")

pilot_survey <- read.csv("C:\\Users\\gelet\\OneDrive\\Documents\\SDS315\\HW3\\nbc_pilotsurvey.csv")

ebay <- read.csv("C:\\Users\\gelet\\OneDrive\\Documents\\SDS315\\HW3\\ebay.csv")
```


# Problem 1

## Theory A

### Claim 

Theory A states that "gas stations will charge more if they lack direct competition in sight". 

### Evidence

```{r echo=FALSE,message=FALSE,warning=FALSE}
# First, generating a simple boxplot to look at differences in the distribution of prices in the sampled gas stations
ggplot(gas_prices, aes(x = factor(Competitors), y = Price)) +
  geom_boxplot(fill = "lightblue1", color = "black") +
  labs(x = "Other gas stations in sight?",
       y = "Price of regular unleaded gas on Apr. 3 2016",
       title = "Sampled Gas Station Price Distributions by Presence of Nearby Competitors")
```
Across the 101 gas stations in the sample, gas stations with no competitors in sight had their median gas price on April 3 2016 about `r median(gas_prices$Price[gas_prices$Competitors == "N"]) - median(gas_prices$Price[gas_prices$Competitors == "Y"])` dollars higher than gas stations with competitors in sight. 

```{r echo=FALSE,message=FALSE,warning=FALSE}
# Setting seed
set.seed(800)

# Difference in means of bootstrap samples
boot_competitors <- do(10000)*diffmean(Price ~ Competitors, data = mosaic::resample( gas_prices))

# Displaying the distribution of difference in mean gas prices between gas stations w/ and w/o competitors across 10,000 bootstrapped samples
ggplot(boot_competitors, aes(x = diffmean)) +
  geom_histogram(fill = "plum1", color = "black") +
  labs(x = "Difference in Mean Gas Prices Between Gas Stations With and Without Competitors",
       y = "Number of Bootstrapped Samples",
       title = "Distribution of Diff. of Bootstrap Sample Mean Gas Prices Between Gas Stations")

# Calculating 95% confidence interval of difference in bootstrapped means
confint(boot_competitors, level = 0.95)
```

10000 bootstrapped samples were constructed from the original sample of 101 gas stations, and in each bootstrap sample the difference in mean gas station prices between gas stations without and with competitors in sight was taken. The distribution of the differences in means is displayed in the histogram above. A 95% confidence interval of the difference in means was calculated; based on this, with 95% confidence, the difference in mean prices between gas stations without and with competitors in sight is between -0.0553 and 0.00853 dollars. 

### Conclusion

The 95% confidence interval of difference in mean gas prices between gas stations without and with competitors captures 0, so the difference in means is not statistically significant at the 5% level. Therefore, there is not enough evidence to show that gas stations without nearby gas stations in sight charge higher gas prices than gas stations with nearby gas stations in sight.

## Theory B

### Claim 

Theory B claims that richer areas have higher gas prices. 

### Evidence

First, it would be helpful to present a scatterplot demonstrating the relationship between gas prices and income within the sample of 101 gas stations.

```{r echo=FALSE,message=FALSE,warning=FALSE}
ggplot(gas_prices, aes(x = Income, y = Price)) +
  geom_point(size = 2) +
  # Line fitting
  geom_smooth(method = lm, se = T)
  labs(x = "2014 Median Household Income by Zip Code of Gas Stations",
       y = "Price of Regular Unleaded Gas on Apr. 3 2016",
       title = "Relationship Between Income and Gas Prices across 101 Sampled Gas Stations")

# Linear Model
summary(lm(data = gas_prices, formula = Price ~ Income))
```

There appears to be a moderate, positive relationship between income and gas prices among the 101 sampled gas stations, as indicated by a Pearson correlation of `r round(cor(gas_prices$Price, gas_prices$Income), 4)`.

```{r echo=FALSE,message=FALSE,warning=FALSE}
# Generating bootstrap samples
bootstrap_income <- do(10000)*lm(data = mosaic::resample(gas_prices), formula = Price ~ Income)

# Visualizing distribution of slope
ggplot(bootstrap_income, aes(x = Income)) +
  geom_histogram(fill = "plum1", color = "black") +
  labs(x = "Bootstrapped Mean Change in Gas Price for Every $1 Increase in Income",
       y = "Number of Bootstrapped Samples",
       title = "Bootstrapped Linear Regression Slope Coefficients of Income on Gas Prices")

# Confidence Interval Calculation
confint(bootstrap_income, level = 0.95)
```
10000 bootstrapped samples were taken and a linear regression output was generated for each one, with income as the predictor and gas prices as the outcome variable. Above is the distribution of the bootstrapped linear regression slopes. A confidence interval was calculated from the bootstrapped samples, finding with 95% confidence that the mean change in gas prices for every one dollar increase in income lies between 0.0000006513 and 0.000001787764 dollars. 

### Conclusion

Since the 95% confidence interval of mean change of gas prices for every dollar increase in income does not capture 0, the data do appear to support the theory that richer areas have higher gas prices.

## Theory C

### Claim

Theory C states that gas stations at stoplights have higher gas prices than gas stations not at stoplights.

### Evidence

```{r echo=FALSE,message=FALSE,warning=FALSE}
# Visualizing difference in mean prices between sampled gas stations at and not at stoplights
ggplot(gas_prices, aes(x = factor(Stoplight), y = Price)) +
  geom_boxplot(fill = "lightblue1", color = "black") +
  labs(x = "Stoplight In Front of Gas Station?", 
       y = "Price of Regular Unleaded Gasoline on Apr. 3 2016",
       title = "Differences in Gas Prices Across Sampled Gas Stations by Presence of Stoplight")

round(mean(gas_prices$Price[gas_prices$Stoplight == "Y"]) - mean(gas_prices$Price[gas_prices$Stoplight == "N"]), 4)
```

Sampled gas stations with a stoplight in front of them had a mean gas price around `r round(mean(gas_prices$Price[gas_prices$Stoplight == "N"]) - mean(gas_prices$Price[gas_prices$Stoplight == "Y"]), 4)` dollars lower than gas stations without stoplights in front of them.

```{r echo=FALSE,message=FALSE,warning=FALSE}
# Generating bootstrapped diffs. in means and storing in object bootstrap_stoplights
bootstrap_stoplights <- do(10000)*diffmean(Price ~ Stoplight, data = mosaic::resample(gas_prices))

# Visualizing distribution of bootstrapped diff. in means
ggplot(bootstrap_stoplights, aes(x = diffmean)) +
  geom_histogram(fill = "plum1", color = "black") +
  labs(x = "Bootstrapped Difference in Mean Gas Price By Presence of Stoplight",
       y = "Number of Bootstrapped Samples",
       title = "Distribution of Bootstrapped Diff. of Mean Gas Prices Between Stoplight and Non-Stoplight Stations")

# Calculating a confidence interval of the diff. in means
confint(bootstrap_stoplights, level = 0.95)
```

Bootstrap samples were generated from the original sample, and for each the difference in mean gas prices between gas stations with and without stoplights in front of them was calculated; this process was repeated 10,000 times and the output was stored in an object and the distribution of the bootstrapped diff. in means is shown in the histogram above. A confidence interval was generated from the 10000 bootstrapped diff. in means, determining with 95% confidence that the difference in mean gas prices between gas stations with and without a stoplight in front of them lies between -0.0378 and 0.0302 dollars. 

### Conclusion

Since the 95% confidence interval of the difference in bootstrapped gas price means between gas statiosn with and without a stoplight in front of them captures 0, the data do not support the theory that gas stations with a stoplight in front of them have higher gas prices.

## Theory D

### Claim

Theory D claims that gas stations accessible from a highway have higher gas prices than gas stations that are not accessible from a highway.

### Evidence

```{r echo=FALSE,message=FALSE,warning=FALSE}
# Visualizing the difference in gas prices between gas stations accessible and not accessible from a highway
ggplot(gas_prices, aes(x = factor(Highway), y = Price)) +
  geom_boxplot(fill = "lightblue1", color = "black") +
  labs(x = "Gas Station Accessible from Highway or Highway Access Road?",
       y = "Price of Regular Unleaded Gasoline on Apr. 3 2016",
       title = "Sampled Gas Station Gas Prices by Highway Accessibility") 

round(mean(gas_prices$Price[gas_prices$Highway == "Y"]) - mean(gas_prices$Price[gas_prices$Highway == "N"]), 4)
```
Sampled gas stations that are accessible from a highway have gas prices `r round(mean(gas_prices$Price[gas_prices$Highway == "Y"]) - mean(gas_prices$Price[gas_prices$Highway == "N"]), 4)` dollars higher than those unaccessible from a highway.

```{r echo=FALSE,message=FALSE,warning=FALSE}
# Generating bootstrap diff in means between gas stations accessible and unaccessible from highway
bootstrap_highway <- do(10000)*diffmean(Price ~ Highway, data = mosaic::resample(gas_prices))

# Visualizing bootstrap diff in means
ggplot(bootstrap_highway, aes(x = diffmean)) +
  geom_histogram(fill = "plum1", color = "black") +
  labs(x = "Bootstrapped Difference in Mean Gas Price",
       y = "Number of Bootstrapped Samples",
       title = "Distribution of Bootstrapped Diff. in Mean Gas Prices by Highway Accessibility to Gas Station")

# Calculating confidence interval from diff. in means
confint(bootstrap_highway, level = 0.95)
```
Bootstrap samples were taken from the original sample, and the difference in mean gas prices between gas stations with and without highway accessibility was calculated; this process was repeated 10000 times and the output is displayed in the histogram distribution above. A confidence interval was calculated on the diff. in means, finding with 95% confidence that the difference in mean gas prices between gas stations with and without highway or highway access road accessibility lies between 0.00883 and 0.08134 dollars. 

### Conclusion

Since the 95% confidence interval of the difference in mean gas prices betwen gas stations with and without highway or highway access road accessbility does not capture 0, the data do support the theory that gas stations with direct highway access have higher gas prices.

## Theory E

### Claim

Theory E claims that Shell brand gas stations have higher gas prices than non Shell brand gas stations.

### Evidence 

```{r echo=FALSE,message=FALSE,warning=FALSE}
# Creating a binary value where Y indicates Shell-brand, N indicates non-Shell Brand
gas_prices <- gas_prices %>% 
  mutate(Shell = ifelse(Brand == "Shell", "Y", "N"))

# Visualizing difference in gas prices among sampled gas stations by brand
ggplot(gas_prices, aes(x = factor(Shell), y = Price)) +
  geom_boxplot(fill = "lightblue1", color = "black") +
  labs(x = "Shell Gas Station?", 
       y = "Price of Regular Unleaded Gasoline on Apr. 3 2016",
       title = "Differences in Gas Prices Among Sampled Gas Stations By Shell Brand")
```
Sampled Shell gas stations had a mean gas price `r round(mean(gas_prices$Price[gas_prices$Shell == "Y"]) - mean(gas_prices$Price[gas_prices$Shell == "N"]), 4)` dollars higher than non Shell gas stations.

```{r echo=FALSE,message=FALSE,warning=FALSE}
# Generating 10000 bootstrapped diff. in means 
bootstrap_shell <- do(10000)*diffmean(Price ~ Shell, data = mosaic::resample(gas_prices))

# Visualizing bootstrapped diff. in means
ggplot(bootstrap_shell, aes(x = diffmean)) +
  geom_histogram(fill = "plum1", color = "black") +
  labs(x = "Bootstrapped Difference in Mean Gas Price",
       y = "Number of Bootstrap Samples", 
       title = "Distribution of Bootstrapped Diff. in Mean Gas Prices Between Shell/Non-Shell Gas Stations")

# Creating confidence interval from diff. in means
confint(bootstrap_shell, level = 0.95)
```
Bootstrap samples were taken from the original sample and the difference in mean gas prices between Shell and non-Shell gas stations was calculated for each bootstrap sample, with the process repeated 10000 times and the distribution of difference in means displayed on the graph above. A confidence interval was produced, determinging with 95% confidence that the difference in mean gas prices between Shell and non-Shell gas stations lies between -0.00928 and 0.0653 dollars.

### Conclusion

Since the 95% confidence interval of difference in mean gas prices between Shell and non-Shell gas stations captures 0, the data do not support the theory that Shell gas stations charge higher gas prices than non-Shell gas stations.

# Problem 2

## Part A

```{r echo=FALSE,message=FALSE,warning=FALSE}
# Filtering and creating new dataframe
sclass_2011 <- s_class %>% 
  filter(year == 2011 & trim == "63 AMG")

# Generating bootstrapped samples
boot_sclass_2011 <- do(10000)*mean(~mileage, data = mosaic::resample(sclass_2011)) 

# Calculating confidence interval
confint(boot_sclass_2011, level=0.95)
```

It can be stated with 95% confidence that the mean mileage of 2011 Mercedes S-Class 63 AMGs on cars.com at the time of data collection lies between 26223.03 and 31771.8 miles per gallon.

## Part B

```{r echo=FALSE,message=FALSE,warning=FALSE}
# Filtering, creating variable isBlack and creating new dataframe
sclass_2014 <- s_class %>% 
  filter(year == 2014 & trim == "550") %>% 
  mutate(isBlack = ifelse(color == "Black", yes=TRUE, no=FALSE))

# Generating bootstrapped samples
boot_sclass_2014 <- do(10000)*prop(~isBlack, data = mosaic::resample(sclass_2014))

# Calculating confidence interval
confint(boot_sclass_2014, level = 0.95)
```
It can be stated with 95% confidence that the proportion of 2014 Mercedes S-Class 550s on cars.com at the time of data collection that were black lies between 0.4168 and 0.4534.

# Problem 3

## Part A

### Question

What show makes people happier: "Living with Ed" or "My Name is Earl"?

### Approach

A dataset is created, only including observations on "Living with Ed" or "My Name is Earl",
and to measure happiness, the `Q1_Happy` variable is used. The difference in mean scores in `Q1_Happy` between "Living with Ed" and "My Name is Earl" viewers is computed from a bootstrap sample constructed from the original sample, and this process is repeated 10000 times to generate 10000 bootstrapped differences in mean `Q1_Happy` scores. Then, a 95% confidence interval is constructed from the bootstrapped samples to evaluate the difference in mean `Q1_Happy` scores.

### Results

```{r echo=FALSE,message=FALSE,warning=FALSE}
# Constructing new dataset with Ed and Earl shows
EdEarl_pilotsurvey <- pilot_survey %>% 
  filter(Show == "Living with Ed" | Show == "My Name is Earl")

# Factored boxplot to observe differences in happiness in the original sample
ggplot(EdEarl_pilotsurvey, aes(x = factor(Show), y = Q1_Happy)) +
  geom_boxplot(fill = "lightblue1", color = "black") +
  labs(x = "Show",
       y = "Happiness Score",
       title = "NBC Pilot Survey Happiness Scores Between Ed and Earl Shows")
```

The graph above displays differences in the distribution of happiness scores between "My Name is Earl" and "Living with Ed" in the survey sample. Both shows have a median happiness score of 4, but "Living with Ed" appears to have a wider range of score values than "My Name is Earl".

```{r echo=FALSE,message=FALSE,warning=FALSE}
# Constructing 10000 bootstrapped difference in means
boot_EdnEarl <- do(10000)*diffmean(Q1_Happy ~ Show, data = mosaic::resample(EdEarl_pilotsurvey))

# Visualization of distribution of bootstrapped differences of means
ggplot(boot_EdnEarl, aes(x = diffmean)) +
  geom_histogram(fill = "plum1", color = "black") +
  labs(x = "Bootstrapped Difference in Mean Happiness Scores",
       y = "Number of Bootstrap Samples",
       title = "Bootstrapped Diff. in Mean Happiness Scores Between 'Ed' and 'Earl' Shows")

# Constructing 95% confidence interval for bootstrapped diff. in means
confint(boot_EdnEarl, level = 0.95)
```
The graph above displays the distribution of the bootstrapped differences in mean happiness scores.

### Conclusion

The confidence interval for these differences in means demonstrates with 95% confidence that the difference in mean happiness scores between "Living with Ed" and "My Name is Earl" is between -0.3991 and 0.1004. However, since 0 is within this range (a difference of means equal to 0 would mean there is no difference in happiness scores between the two shows), the data do not support the theory that viewers feel different levels of happiness when watching "Living with Ed" and "My Name is Earl".

## Part B

### Question

We want to answer: which shows made viewers more annoyed, "The Biggest Loser" or "The Apprentince: Los Angeles"?

### Approach

A dataset is created only including the shows "The Biggest Loser" and "The Apprentince: Los Angeles", and differences in the values of `Q1_Annoyed` in the original sample are visualized using a factored boxplot. Then, 10000 bootstrapped difference in mean annoyance scores are generated from the original dataset, visualized on a histogram. Finally, a 95% confidence interval is constructed from the bootstrapped difference in mean annoyance scores to examine the strength of evidence that one show causes more annoyance than the other.

### Results

```{r echo=FALSE,message=FALSE,warning=FALSE}
# Filtering dataset to only include Loser and Apprentince and creating new dataframe
LoserApprn_pilotsurvey <- pilot_survey %>% 
  filter(Show == "The Biggest Loser" | Show == "The Apprentice: Los Angeles")

# Visualizing sample differences
ggplot(LoserApprn_pilotsurvey, aes(x = factor(Show), y = Q1_Annoyed)) +
  geom_boxplot(fill = "lightblue1", color = "black") +
  labs(x = "Show",
       y = "Annoyance Score",
       title = "NBC Pilot Survey Annoyance Scores Between 'The Biggest Loser' and 'The Apprentice'")
```
Within the NBC sample, "The Apprentince: Los Angeles" and "The Biggest Loser" appear to have identical annoyance score distributions, with median annoyance scores of 2.

```{r echo=FALSE,message=FALSE,warning=FALSE}
# Generating bootstrapped samples
boot_LoserAppren <- do(10000)*diffmean(Q1_Annoyed ~ Show, data = mosaic::resample(LoserApprn_pilotsurvey))

# Visualization of distribution of bootstrapped differences of means
ggplot(boot_LoserAppren, aes(x = diffmean)) +
  geom_histogram(fill = "plum1", color = "black") +
  labs(x = "Bootstrapped Difference in Mean Annoyance Scores",
       y = "Number of Bootstrap Samples",
       title = "Bootstrapped Diff. in Mean Annoyance Scores Between 'Loser' and 'Apprentice' Shows")

# Constructing 95% confidence interval
confint(boot_LoserAppren, level = 0.95)
```
The graph above displays the bootstrapped difference in mean annoyance scores between "The Biggest Loser" and "The Apprentice: Los Angeles". 

### Conclusion

The confidence interval computed demonstrates with 95% confidence that the difference in mean annoyance scores between "The Biggest Loser" and "The Apprentice: Los Angeles" lies between -0.5231 and -0.0218. Since 0 is not within this interval, the data do support the idea that viewers had different levels of annoyance between "The Biggest Loser" and "The Apprentice: Los Angeles". 

## Part C

### Question

What is the proportion of American TV watchers expected to have given a score of 4 or above to the NBC's Q2_Confusing pilot survey question for the show "Dancing with the Stars"?

### Approach

A dataset is created from the original survey only containing observations with the show "Dancing with the Stars", and a new binary variable is created that assigns a TRUE or FALSE value depending on whether or not a viewer gave a confusion score of 4 or higher. 10000 bootstrap samples are generated, and for each the proportion of viewers who gave "Dancing with the Stars" a confusion score of 4 or higher is calculated and stored as entries into a dataframe. Then, a 95% confidence interval is constructed from the bootstrapped proportions.

### Results

```{r echo=FALSE,message=FALSE,warning=FALSE}
dancing_pilotsurvey <- pilot_survey %>% 
  # Filtering dataset to only include Dancing with the Stars
  filter(Show == "Dancing with the Stars") %>% 
  # Creating a binary variable for confusion scores 4+
  mutate(isConfusedFour = ifelse(Q2_Confusing >= 4, yes=TRUE, no=FALSE))

# Generating bootstrap sample proportions
boot_dancing <- do(10000)*prop(~isConfusedFour, data = mosaic::resample(dancing_pilotsurvey))

# Visualizing bootstrapped proportions
ggplot(boot_dancing, aes(x = prop_TRUE)) +
  geom_histogram(fill = "plum1", color = "black") +
  labs(x = "Bootstrap Proportion of Viewers Giving Confusion Score 4+",
       y = "Number of Bootstrap Samples",
       title = "Distribution of Bootstrapped Proportions of Confusion Scores 4+ for Dancing w/the Stars")

# Generating 95% confidence interval
confint(boot_dancing, level = 0.95)
```

Above is the distribution of the bootstrapped proportions of viewers giving "Dancing with the Stars" a confusion score of 4 or greater. It is centered at around 0.07-0.08.

### Conclusion 

The confidence interval generated determines with 95% confidence that the proportion of American TV watchers expected to have given a score of 4 or above to the NBC's Q2_Confusing pilot survey question for the show "Dancing with the Stars" lies between 0.03887 and 0.11602.

# Problem 4

## Question

Does Ebay's paid advertising on Google AdWords increase revenue for Ebay?

## Approach

To determine whether revenue changed before to after the experiment, a revenue ratio variable is created; that is, the ratio of revenue 30 days after the experiment to revenue 30 days before the experiment. 10000 bootstrap samples are taken, with the difference in the mean revenue ratio between treatment and control DMAs calculated for each sample. The distribution of these differences in means is visualized in a histogram, and a 95% confidence interval is constructed for the difference in mean revenue ratio.

## Results

```{r echo=FALSE,message=FALSE,warning=FALSE}
# Creating revenue ratio variable
ebay <- ebay %>% 
  mutate(rev_ratio = rev_after/rev_before)

# Distributions of revenue ratio between treatment and control DMAs in sample
ggplot(ebay, aes(x = factor(adwords_pause), y = rev_ratio)) +
  geom_boxplot(fill = "lightblue1", color = "black") +
  labs(x = "DMA Group (Treatment = 1, Control = 0)",
       y = "Revenue Ratio",
       title = "Distribution of Revenue Ratios in Ebay Adwords Experiment By Group Type")
```

The factored boxplot above shows the distribution of revenue ratios across DMAs in Ebay's Adwords experiment by the DMA's experimental group, with treatment groups that experienced a paid ad pause designated by 1, and control groups that had no paid ad pause indicated by 0. DMAs in the treatment group appeared to have lower revenue ratios than control DMAs, indicated that the former saw relatively lower revenue after the experiment. 

```{r echo=FALSE,message=FALSE,warning=FALSE}
# Generating bootstrap samples
boot_ebay <- do(10000)*diffmean(rev_ratio ~ adwords_pause, data = mosaic::resample(ebay))

# Visualizing bootstrapped differences in mean revenue ratios
ggplot(boot_ebay, aes(x = diffmean)) +
  geom_histogram(fill = "plum1", color = "black") +
  labs(x = "Bootstrapped Difference in Mean Revenue Ratio",
       y = "Number of Bootstrap Samples",
       title = "Bootstrapped Differences in Mean Revenue Ratios Between Treatment and Control")

# Constructing 95% confidence interval
confint(boot_ebay, level = 0.95)
```

The graph above displays the distribution of 10000 bootstrapped differences in mean revenue ratios.

## Conclusion

The confidence interval computed determines with 95% confidence that the difference in the mean revenue ratio between DMAs that had paid ads on Google AdWOrds paused for a month and DMAs that continued running paid ads on Google Adwords lies between -0.09115 and -0.01396. In other terms, it appears that DMAs that paused paid Google ads saw generally lower revenue ratios than DMAs that did not pause paid Google ads. Therefore, the data provide support for the theory that Ebay's paid advertising on Google AdWords increases revenue for Ebay.  